### 2016-09-16 paper/conclusion.tex
added rough draft of all sections

Conclusion

>>>>

We have coded up an implementation of type directed synthesis on a domain
specific language of bijective lenses.  This program allows users to input two
regular expressions, and a few input/output examples, and synthesizes a
bijective lens between those regular expressions, matching those input/output
examples.  While lenses are very useful for their guarantees, they are a
difficult domain specific language to code in, in practice.  This program
provides an easier interaction model for the creation of lenses through using
synthesis as the way to interact with the programmer.

This synthesis algorithm is competitive with state of the art tools for
synthesis on strings transformations.  This gain is made through requiring a
larger amount of user interaction with the synthesis algorithm.  However, in the
for this increased interaction, the program provides the capabilities of
synthesizing very complicated transformations, and provides stronger guarantees
about correctness on inputs dissimilar to those provided in the examples.

### 2016-09-16 paper/discussion.tex
added variable information on master file to each file
added rough draft of all sections

\section{Conclusion}

Conclusion

>>>>

\section{Discussion}

We found that in practice, requiring a user to put in more information for the
synthesis of a string transformation creates a more reliable algorithm.
However, this comes at a few costs.  The user needs to know how to write out
regular expressions.

The largest difficulty in this is the restriction to bijective lenses.
Oftentimes, we find the need to hide information away in some alternate form.
For example, there is oftentimes whitespace in the definition of a regular
expression.  This whitespace must be copied somehow over to the other side.
Oftentimes this requires the use of having a portion of the string dedicated to
holding the whitespace information of the other side.

Another annoyance is the fact that if the correct lens does not exist,
rarely does the tool terminate.  Much of the time it will just spin trying to
perform increasingly complicated transformations to find the correct lens.
A better solution would be to find something close to the correct lens, and have
a user interaction model that supports iterative synthesis.  The user could give
more examples which fix where the program is wrong.

\subsection{Future Work}
Some of the issues seen in this have already been observed.  For example, in the
quotient lens paper, the issue of having to hack a place in the source string to
move whitespace information from the view to the source was noted.  An extension
we would like to work on would be able to synthesize quotient lenses.

Another approach to having disparate information on each side of the lens is
through the use of symmetric lenses.  This could be another approach to the
whitespace problem.

### 2016-09-16 paper/evaluation.tex
added variable information on master file to each file
added rough draft of all sections

\section{Evaluation}
We ran our algorithm on tests validating our ability to handle regular
expression equivalences, tests derived from the Augeus test suite, and from
tests derived from the FlashFill tests.\bcp{Need to carefully explain our
  experimental setup and methodology (on what machine config, how many test
  runs, what kind of statistics, ...)}

\input{fig_runtimes-table}
\subsection{Speed}
Figure~\ref{fig:runtimes-table} provides the runtime of our algorithm on our
test suite.
In this figure, we want to minimize runtime, as a smaller runtime means we were
able to synthesize very efficiently.
We found that, on many of our difficult examples, our algorithm was able to find
the appropriate lens very quickly.

@@

\input{fig_examples-required-table}
As we see in Figure~\ref{fig:examples-required-table}, examples are not too
important for synthesizing bijective lenses.\bcp{If that's really the
  conclusion, then maybe we should have written a paper about synthesizing
  bijective lenses {\em without} examples!  But I don't think we can
  actually conclude this from the examples we have.  }
Because bijective lenses have very few well typed programs, compared to the
number of functions, oftentimes just a few examples is sufficient.
As would be expected, the functions that have to make choices about the
locations of where data goes are those that require more examples.
In particular, the address problem requires more examples, because it must make
choices of where information goes, like which name is the a person's first name,
and which is their last.

\input{fig_expanded-time-table}
\subsection{Importance of User Defined Data Types}

>>>>

\section{Evaluation}
We ran our algorithm on tests validating our ability to handle regular
expression equivalences, tests derived from the Augeus test suite, and from
tests derived from the FlashFill tests.
The values presented are the arithmetic means of 10 test runs.
The tests were run on a 2.5 GHz Intel Core i7 processor, with
16 GB of 1600 MHz DDR3, running Mac OS X Yosemite.

\begin{figure*}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\bfseries Test & \bfseries ForceExpandTime & \bfseries ForceExpandExamplesRequired
& \bfseries Computation Time & \bfseries Examples Required
\csvreader[head to column names]{generated-data/data.csv}{}
{\\\hline\Test & \ForceExpandTime & \ForceExpandExamplesRequired & \ComputationTime & \ExamplesRequired}
\\\hline
\end{tabular}

\end{figure*}

\subsection{Speed}
Figure~\ref{fig:runtimes-table} provides the runtime of our algorithm on our
test suite.
In this figure, we want to minimize runtime, as a smaller runtime means we were
able to synthesize very efficiently.
We found that, on many of our difficult examples, our algorithm was able to find
the appropriate lens very quickly.

@@

As we see in Figure~\ref{fig:examples-required-table}, examples are not too
important for synthesizing bijective lenses.\bcp{If that's really the
  conclusion, then maybe we should have written a paper about synthesizing
  bijective lenses {\em without\/} examples!  But I don't think we can
  actually conclude this from the examples we have.  }
Because bijective lenses have very few well typed programs, compared to the
number of functions, oftentimes just a few examples is sufficient.
As would be expected, the functions that have to make choices about the
locations of where data goes are those that require more examples.
In particular, the address problem requires more examples, because it must make
choices of where information goes, like which name is the a person's first name,
and which is their last.

\subsection{Importance of User Defined Data Types}

User defined data types are very critical to the algorithm and performance of
the synthesis algorithm.  User defined data types help direct the search of
which expansions should be performed.  Furthermore, they help to reduce the
number of clauses we have to deal with.  There is a potentially exponentially
large blowup in converting regular expressions into DNF form, but keeping
certain portions of the regular expressions abstracted lessens that blowup,
as those regular expressions are kept atomic in converting into DNF form.

User defined data types are also critical in the applications of previously
defined lenses.  We restrict the application of previously defined lenses to
only on user defined data types in the synthesis algorithm.  We then group the
user defined data types based on which have previously defined lenses between
them.  This allows for efficient recognition of when two user defined regular
expressions can be mapped to each other.  Removing the ability to use previously
defined lenses makes certain problems intractable, like (example1, example2,
example3).

While we can simulate the use of user defined data types automatically, by
giving regular expressions which have the same syntax the same user defined data
type, manually inputting them performs better in practice.  This is because the
procedure for automatically creating user defined data types creates far more
data types, which then creates a more complicated search problem.

%%% Local Variables:
%%% TeX-master: "main"
%%% End:

### 2016-09-16 paper/extended-example.tex
added variable information on master file to each file
added rough draft of all sections

\dpw{the lstlisting environment is drawing a line underneath the code.
I don't want that.  we need to fix it but I'm focusing on writing.
Also, it may be good to hack it so it can be in a backslash small font.
There is perhaps another environment we could use -- a code environment.}

@@

Figure~\cite{fig:bibend-spec} presents
the user's specification of such a problem.  In general, such specifications are
given by defining the source and target data formats~\footnote{Lenses
operate both left-to-right and right-to-left, so both formats are sometimes
sources and sometimes targets. Nevertheless, we will often call the
format on the left, the ``source,'' and the format on the right, 
the ``target.''} as well as some examples.  In this case, the lens
we wish to synthesize is \CF{BibEnd}.  It converts between 
the formats \CF{Bib} and \CF{End}.
Such formats are defined using a standard syntax for regular expressions.
Intermediate regular expressions are named in order to facilitate
construction of complex formats.

@@

The simplest lens in the combinator language is the identity lens 
at type $R$,
$\IdentityLensOf{R}$, which copies data matching regular expression $R$
from source to target, and vice versa.  The
constant lens $\ConstLens{\String_1}{\String_2}$, when operated left-to-right, 
replaces string $\String_1$ with $\String_2$, and when operated right-to-left,
replaces string $\String_2$ with $\String_1$.

Each of the other lenses manipulate structured data.  For instance,
$\ConcatLens{\Lens_1}{\Lens_2}$

@@

There is no lens that goes between \BibTex{} and \EndNote{}, so one of them must
be expanded.
Indeed, both of them must be expanded, as \BibTex{} will never be able to be
the source of any lens in this problem, nor will \EndNote{} be able to be the
target.
This can be generalized for the sub-expressions of \BibTex{} and \EndNote{},  
If, any user defined data type does not exist in the possible types of the
other side, it must be expanded.

FIGURE:
Trees of user defined regular expressions, with red x's over those that must be
removed.

After completing these necessary expansions, the problem becomes synthesizing a
lens between the regular expression
\begin{lstlisting}
Bib: ""   | ("author={" Name ", " Name Names (" and " (" Name ", " Name Names))* "},")

End: ("%A " (Name Names " " Name) "\n")*
\end{lstlisting}
and

A basic synthesis approach from here will quickly fail.  Both \texttt{BibTex'}
and \texttt{EndNote'} are of the form $\Regex_1\Concat\Regex_2$, but there is
not a lens between \texttt{"\%0"} and
\texttt{("@inproceedings" | "@article" | "@book")}.

These issues can be handled by normalizing the regular expressions slightly.
By putting them in a form that normalizes away the equivalences of
associativity, distributivity, and \EmptyString{} identity.

The normalized form of these equivalences is too long to be contained in this
paper.  Instead, we will work through \afm{or should we just do the easier
example shown above}.

With those equivalences are handled, perhaps the lens can be synthesized.
A lens can be synthesized, if there is some pairing of sequences such that
there are sequence lenses between the two elements of the pair for all pairings.
A sequence lens can be synthesized if there is some pairing of atoms, such that
there is an atom lens between the two atoms.  An atom lens can be synthesized
if the regular expressions the same user defined data type, or if there is a dnf
lens between the two dnfs under the stars.

FIGURE:

"" | Author Author*

AuthorEndNote*

Looking at the number of sequences, we can easily see that these dnf regular
expressions do not have a lens between them, without some rewrite rules applied.
However, there are a large number of possible rewrite rules that can be applied.
Instead of taking and testing all of them, each rewrite rule is applied, and
then placed in a priority queue.
The priority for each problem is a linear combination of the number of rewrites
applied, and the distance between the two dnf regular expressions by a
pseudometric.

This pseudometric is defined as the sum of two metrics.
The first metric is a standard distance metric over the size of the two dnf
regular expressions.
The second metric is defined based on the distribution of user defined regular
expressions.
We can create an infinite dimensional free module over the integers from the
regular expressions, generated by the set $\SetOf{(U,i)\SuchThat U\text{is a
user defined data type}\BooleanAnd i\in\Nats}$.
We can provide a function $f$ that maps all regular expressions into the element
which counts how many user defined data types there are at a given depth.
\afm{i should probably write down this function with a mapsto style, instead of
just through text}
The second metric is the taxicab metric on the elements of the vector space
mapped to by the regular expressions.

FIGURE:

\begin{lstlisting}
"" | (name,name) (name,name)* -> \{ ((Author,0),2) , ((Author,1),4) \}
(name,name)* -> \{ ((name,1),2) \}

d("" | (name,name) (name,name)*,(name,name)*) =
taxicab(\{ ((Author,0),2) , ((Author,1),4) \},other)+(8-2)
\end{lstlisting}

Empirically, good transformations, like (name,name)* -> (name,name)+
do very well, giving a distance of zero, where bad transformations, like
name -> [A-Z][a-z]* do not do as well.

Using the priority queue, the problem that gets popped would be
"" | (name,name) (name,name)* <=> "" | (name,name) (name,name)*.
Unlike the previous example, this cannot be so easily determined to not be
equivalent.

A naive approach to this would be to attempt all possible pairings of sequences,
and all pairings of atoms within those sequences.  However, the complexity of
trying all permutations is (\# permutations)!.

Instead, an ordering can be provided to atoms, clauses, and dnf regular
expressions, where two elements are equal in the ordering if there exists a lens
between them.  Upon applying that ordering to the elements of the dnf regular
expression, determining whether there is a lens between two dnf regular
expressions amounts to merely seeing if each element of the ordering is
equal with respect to the ordering, as the one on the other side.

FIGURE
\begin{lstlisting}
e | (n,n)(n,n)* <=> e | (n,n)(n,n)*
reorder
e | (n,n)(n,n)* <=> e | (n,n)(n,n)*
matched together
e <=> e , (n,n) <=> (n,n) , etc.
\end{lstlisting}

This still isn't complete yet, for we don't know if the generated lens will
satisfy the examples.  The same problem arises with multiple permutations
being possible.  If all the sequences can be mapped to each other, then there is
a valid lens for each matching of the sequences!  And not all of these
valid lenses satisfy the examples.
To resolve this issue, we embed the parse trees of the examples into the
regular expressions, and treat regular expressions with different aspects of the
parse trees embedded as differently in the ordering.

FIGURE
e | name name(name, name)* , lalalaexample

\begin{lstlisting}
 /\
e  |
   /\
 name *
      .
     /\
   name name

lalaexample -> or2 (lala . Iterated(ex.am,p.le)

combine to

(the top thing but with the parse tree embedded alongside the expressions)
\end{lstlisting}

>>>>

\dpw{It may be good to hack lstlisting so it can be in a backslash small font.
There is perhaps another environment we could use --- a code environment.}

@@

Figure~\ref{fig:bibend-spec} presents
the user's specification of such a problem.  In general, such specifications are
given by defining the source and target data formats~\footnote{Lenses
operate both left-to-right and right-to-left, so both formats are sometimes
sources and sometimes targets. Nevertheless, we will often call the
format on the left, the ``source,'' and the format on the right, 
the ``target.''} as well as some examples.  In this case, the lens
we wish to synthesize is \CF{BibEnd}.  It converts between 
the formats \CF{Bib} and \CF{End}.
Such formats are defined using a standard syntax for regular expressions.
Intermediate regular expressions are named in order to facilitate
construction of complex formats.

@@

The simplest lens in the combinator language is the identity lens 
at type $R$,
$\IdentityLensOf{R}$, which copies data matching regular expression $R$
from source to target, and vice versa.  The
constant lens $\ConstLensOf{\String_1}{\String_2}$, when operated left-to-right, 
replaces string $\String_1$ with $\String_2$, and when operated right-to-left,
replaces string $\String_2$ with $\String_1$.

Each of the other lenses manipulate structured data.  For instance,
$\ConcatLensOf{\Lens_1}{\Lens_2}$

@@

There is no lens that goes between \BibTex{} and \EndNote{}, so one of them must
be expanded.
Indeed, both of them must be expanded, as \BibTex{} will never be able to be
the source of any lens in this problem, nor will \EndNote{} be able to be the
target.
This can be generalized for the sub-expressions of \BibTex{} and \EndNote{},  
If, any user defined data type does not exist in the possible types of the
other side, it must be expanded, shown in Figure~\ref{fig:expanded-defs}.

\begin{figure}
  \Tree[.\textcolor{dkred}{\sout{\texttt{Bib}}}
    [.\textcolor{dkred}{\sout{\texttt{BNames}}}
      \texttt{Name}
      [.\texttt{Names}
        \texttt{Name} ]]]
  \Tree[.\textcolor{dkred}{\sout{\texttt{End}}}
    [.\textcolor{dkred}{\sout{\texttt{ENames}}}
      \texttt{Name}
      [.\texttt{Names}
        \texttt{Name} ]]]
\caption{Marking of Definitions which must be expanded}
\label{fig:expanded-defs}
\end{figure}

Trees of user defined regular expressions,
with those that must be replaced by their definitions crossed out.

After completing these necessary expansions, the problem becomes synthesizing a
lens between the regular expression
\begin{lstlisting}
Bib': ""   | ("author={" Name ", " Name Names (" and " (Name ", " Name Names))* "},")

End': ("%A " (Name Names " " Name) "\n")*
\end{lstlisting}
and

A basic type directed synthesis approach from here will quickly fail.
Both \texttt{Bib'}
and \texttt{End'} are of the form $\Regex_1\Concat\Regex_2$, but there is
not a lens between \texttt{"\%0"} and
\texttt{("@inproceedings" | "@article" | "@book")}.

These issues can be handled by normalizing the regular expressions slightly.
By putting them in a form that normalizes away the equivalences of
associativity, distributivity, and \EmptyString{}.

\begin{lstlisting}
DNFBib':
  ""
  | "author=" Name ", " Name "" Names ""
    (" and " Name ", " Name "" Names "")* ""

DNFEnd':
  "" ("%A " Name "" Names " " Name "\n")* ""
\end{lstlisting}

With those equivalences handled, one might think it is now possible to
synthesize the lens.
Unfortunately, we are still unable to preform any synthesis because the types
on the left and right are quite different, and the lens connectives require
similar types of the left and right.
For example, as \texttt{DNFBib'} has an outer |, we know an or lens must be
synthesized.
However, an or lens requires an outer | on both sides, and \texttt{DNFEnd'} does
not have any outer |.

While, with normalization, certain equivalences have been handled, there are
still many equivalences that haven't been handled.
For example, the star in \texttt{End'} can be unrolled, turning \texttt{End'},
and \texttt{DNFEnd'} into

\begin{lstlisting}
End'': "" | ("%A " (Name Names " " Name) "\n"
("%A " (Name Names " " Name) "\n")*)

DNFEnd'':
  ""
  | "%A " Name "" Names " " Name "\n"
    ("%A " Name "" Names " " Name "\n")* ""
\end{lstlisting}

\texttt{DNFBib'} and \texttt{DNFEnd'} now look very similar.
They both have an outer |, which has on one side a string, and on the other
side, two instances of \texttt{Name}, and an instance of \texttt{Names} outside
of any stars,
with two instances of \texttt{Name}, and an instance of \texttt{Name} underneath
a star.
From this, a lens can easily be made.

However, there is some element of choice in how the lens can be made.
The outer \texttt{Name}s can each be sent to each other.
and the inner \texttt{Name}s can each be sent to each other, so which
should get sent to which.
This can be determined from the examples.
If the first \texttt{Name} in \texttt{DNFBib'} were sent to the first
\texttt{Name} in \texttt{DNFEnd'}, then \texttt{Conway, John} would
get mapped to \texttt{\%a Conway John}, which is inconsistent with the
examples.
The examples can be used, when this situation arises, to specify which parts
of a regular expression can be mapped to which, when the types alone leave
this ambiguous.

\afm{I am not sure going into the metric and priority queue are worth going
  into,
  and least not here?}
%%% Local Variables:
%%% TeX-master: "main"
%%% End:
